SDM\_the\_search\_for\_powers\_of\_2\_everywhere

### 2018-04-17 Chapters/Introduction.tex
Intro with refs...

Despite its widespread use in a number of distinct areas, the rational choice model has not found to be psychologically plausible, for a number of reasons \citep{plous_psychology_1993}.  One of the reasons is that the chosen alternative depends on how decision-makers initially frame a problem (Kahneman and Tversky, 1979). There has been strong criticism of the rational choice model from the heuristics and biases research program, in which problems are carefully devised to show that one's intuitions generally depart from the expected optima, and are generally inconsistent with what would be expected as rational.  A large number of biases that depart from rational choice have been found (see, for instance, \citet{plous_psychology_1993}), placing strain on the traditional rational actor doctrine.  Yet the heuristics and biases studies are concentrated on carefully devised questionnaires applied mostly to undergraduate students --- not on real world settings with serious skin in the game.

Thus a new field of naturalistic decision-making emerged, in which the focus is centered around real life settings and decisions being made under rapidly changing circumstances.  A number of studies have been conducted, from firefighters to nurses to chess players to military personnel.  One of the most interesting theories to emerge from naturalistic decision-making, the recognition-primed decision model, was devised by Gary Klein and his colleagues (Klein, 1999).

@@

Klein (1999) conducted a series of studies with decision-makers under rapidly changing scenarios. During interviews, when questioned how a specific decision (or course of action) was adopted, decision-makers such as the nurse would proclaim, to Klein's frustration, that they `did not make decisions'.  One experienced firefighter proclaimed `I don't make decisions--I don't remember when I've ever made a decision' (Klein, 1999, p.10).   Decision-makers did not seem to be comparing alternative courses of actions, as classical models would predict.  `It is usually obvious what to do in any given situation' (p.11).  Repeated statements of the sort by different decision-makers led Klein to propose a psychologically plausible model of decision-making which radically departed from the established view of `comparing alternatives and selecting the optimum'.

Klein (1999) proposed a model of recognition-primed decision, in which experienced decision-makers would find themselves immersed in complex situations and rapidly take adequate courses of action.  Decision-makers would rapidly perceive cues from any situation and retrieve from episodic memory similar situations (Tulving 1983), which would bring assessments and diagnoses and plausible courses of action.  Because priming mechanisms are automatic and unconscious (Bargh and Chartrand 1999, Bargh et al. 2001), these decision-makers reported doing `the obvious' action in different situations.  This `obvious' course of action, Klein proposes, is brought from long-term episodic memory by priming mechanisms.  Hence, decision-makers would not be selecting among distinct alternatives, but rather simply performing the automatically-provided action.

Even if the `obvious' action seemed plausible for a theory, another problem remained:  if decision-makers did not compare alternatives, then how could they know that a course of action was good?  In subsequent interviews, evidence emerged that decision-makers would be using the simulation heuristic, proposed by Kahneman and Tversky (1982).  That is, facing a particular situation, experienced decision-makers would be primed towards a particular course of action, to the detriment of most alternative courses of action.  This primed alternative would be `simulated', or `run through', one's mind, and, if found acceptable during the simulation processing, would be acted upon without further deliberation.  If problems emerged during mental simulation, another different course of action would be primed. Thus was born a theory of intuitive decision-making, in which experienced people would not be selecting choices from a vast set of alternatives, but instead `testing' their initially primed predispositions with a simulation heuristic.

@@

\begin{quote}
	Before we did this study, we believed the novices impulsively jumped at the first option they could think of, whereas experts carefully deliberated about the merits of different courses of action.  Now it seemed that it was the experts who could generate a single course of action, while novices needed to compare different approaches. --- Klein (1999, p.21)
\end{quote}

Because priming mechanisms that brought plausible actions to mind are unconscious, people would report having ``done the obvious thing to do''.  Decision-makers would be unable to visualize the cognitive processes underlying their decisions, and would in many cases even believe that they had skills of the `fantastic' variety:  One firefighter demands that his whole crew abandon operations inside a house, just to see it collapse seconds afterward.  A radar operator would `chill' after spotting a new track, and would fire counter missiles against it, based on the `feeling' that it was a hostile missile.  It took over a year for this radar operator, after being interviewed by Klein, to understand the incredibly subtle cues that he was responding to whenever he perceived the new radar track.  Unable to reasonably explain their life-saving, rapid, decisions, both the firefighter and the radar operator thought that they had ESP or other fantastic abilities. Careful probing would show that they were able to unconsciously perceive subtle cues, which primed them towards adequate responses.

Beyond a formalization of this process of memory recall, Sparse Distributed Memory will offer us a plausible, both psychologically and neuroscientifically, path towards artificial intelligence.

>>>>

Despite its widespread use in a number of distinct areas, the rational choice model has not found to be psychologically plausible, for a number of reasons \citep{plous_psychology_1993}.  One of the reasons is that the chosen alternative depends on how decision-makers initially frame a problem \citep{kahneman1979prospect}. There has been strong criticism of the rational choice model from the heuristics and biases research program, in which problems are carefully devised to show that one's intuitions generally depart from the expected optima, and are generally inconsistent with what would be expected as rational.  A large number of biases that depart from rational choice have been found (see, for instance, \citet{plous_psychology_1993}), placing strain on the traditional rational actor doctrine.  Yet the heuristics and biases studies are concentrated on carefully devised questionnaires applied mostly to undergraduate students --- not on real world settings with serious skin in the game.

Thus a new field of naturalistic decision-making emerged, in which the focus is centered around real life settings and decisions being made under rapidly changing circumstances.  A number of studies have been conducted, from firefighters to nurses to chess players to military personnel.  One of the most interesting theories to emerge from naturalistic decision-making, the recognition-primed decision model, was devised by \citet{klein1999sources} and his colleagues.

@@

\citet{klein1999sources} conducted a series of studies with decision-makers under rapidly changing scenarios. During interviews, when questioned how a specific decision (or course of action) was adopted, decision-makers such as the nurse would proclaim, to Klein's frustration, that they `did not make decisions'.  One experienced firefighter proclaimed `I don't make decisions--I don't remember when I've ever made a decision' \citep{klein1999sources}, p.10.   Decision-makers did not seem to be comparing alternative courses of actions, as classical models would predict.  `It is usually obvious what to do in any given situation' (p.11).  Repeated statements of the sort by different decision-makers led Klein to propose a psychologically plausible model of decision-making which radically departed from the established view of `comparing alternatives and selecting the optimum'.

\citet{klein1999sources} proposed a model of \emph{recognition-primed decision}, in which experienced decision-makers would find themselves immersed in complex situations and rapidly take adequate courses of action.  Decision-makers would rapidly perceive cues from any situation and retrieve from episodic memory similar situations, which would bring assessments and diagnoses and plausible courses of action.  Because priming mechanisms are automatic and unconscious \citep{bargh1999unbearable, bargh2001automated}, these decision-makers reported doing `the obvious' action in different situations.  This `obvious' course of action, Klein proposes, is brought from long-term episodic memory by priming mechanisms.  Hence, decision-makers would not be selecting among distinct alternatives, but rather simply performing the automatically-provided action.

Even if the `obvious' action seemed plausible for a theory, another problem remained:  if decision-makers did not compare alternatives, then how could they know that a course of action was good?  In subsequent interviews, evidence emerged that decision-makers would be using the simulation heuristic, proposed by \citet{kahneman1982simulation}.  That is, facing a particular situation, experienced decision-makers would be primed towards a particular course of action, to the detriment of most alternative courses of action.  This primed alternative would be `simulated', or `run through', one's mind, and, if found acceptable during the simulation processing, would be acted upon without further deliberation.  If problems emerged during mental simulation, another different course of action would be primed. Thus was born a theory of intuitive decision-making, in which experienced people would not be selecting choices from a vast set of alternatives, but instead `testing' their initially primed predispositions with a simulation heuristic.

@@

\begin{quote}
	Before we did this study, we believed the novices impulsively jumped at the first option they could think of, whereas experts carefully deliberated about the merits of different courses of action.  Now it seemed that it was the experts who could generate a single course of action, while novices needed to compare different approaches. (p.21)
\end{quote}

Because priming mechanisms that brought plausible actions to mind are unconscious, people would report having ``done the obvious thing to do''.  Decision-makers would be unable to visualize the cognitive processes underlying their decisions, and would in many cases even believe that they had skills of the `fantastic' persuasion:  One firefighter demands that his whole crew abandon operations inside a house, just to see it collapse seconds afterward.  A radar operator would `chill' after spotting a new track, and would fire counter missiles against it, based on the `feeling' that it was a hostile missile.  It took over a year for this radar operator, after being carefully and repeatedly interviewed, to understand the incredibly subtle cues that he was responding to whenever he perceived the new radar track.  Unable to reasonably explain their life-saving, rapid, decisions, both the firefighter and the radar operator thought that they had ESP or other fantastic abilities. Careful probing would show that they were able to unconsciously perceive subtle cues, which primed them towards adequate responses.

Beyond a formalization of this process of memory recall, Sparse Distributed Memory will offer us a plausible, both psychologically and neuroscientifically, path towards artificial intelligence.  We'll need to take a turn here.
